---
title: "Data-607 Week-9 Assignment"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 10/26/2018
* > [GitHub Location for rmd file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week9/SachidVijay.Deshmukh60.rmd)
* > [GitHub Location for pdf file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week9/SachidVijay.Deshmukh60.pdf)
* > [RPubs location of published file](http://rpubs.com/sachid/Data607Assignment9)

##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1] Library Initialization

```{r library initialization}
library(jsonlite)
library(tidyr)
library(dplyr)
library(ggplot2)
```

### 2] Read movie critics info

```{r Read movie critics info}
API.URL = "http://api.nytimes.com/svc/movies/v2/reviews/search.json?api-key=ac5b7916e92c4b21926210fbe19e6434&critics-pick=Y"
Movie.Critics <- fromJSON(API.URL)$results
```

### 3] Preview resulting data frame and column types

```{r preview df}
head(Movie.Critics)
str(Movie.Critics)
```

Resulting DF have 20 rows and 16 columns. Following is the more information on data frame columns

*> display_title : Movie Name
*> mpaa_rating : Movie Rating
*> critics_pick : Whether critics pick or not
*> byline : Name of the critics
*> headline : Headline Info
*> summary_short : Short summary
*> publication_date : Date of publication
*> opening_date : Date of movie opening
*> date_updated : Updated Date
*> link.type : Link Type
*> link.url : Link URL
*> link.suggested_link_text : Suggested link url
*> multimedia.type : Type of multimedia
*> multimedia.src : Source of multimedia
*> multimedia.width : Width of multimedia
*> multimedia.height : Height of multimedia

### 4] Fill null values in rating column

```{r fill rating null}
Movie.Critics$mpaa_rating[Movie.Critics$mpaa_rating == ''] <- 'Not Rated'
```

### 5] Show bar chart of rating type

```{r rating type bar chart}
Movie.Critics %>%
   
    ggplot(aes(x=mpaa_rating, fill=mpaa_rating)) + geom_bar(position = "dodge") +# plot
    ggtitle("Movie Rating bar plot")
```

Top most critics reviews are for not rated movies

### 6] Show bar chart of critics

```{r show bar chart as per critic}

Movie.Critics %>%
   
    ggplot(aes(x=byline, fill=byline)) + geom_bar(position = "dodge") +# plot
    ggtitle("Critic bar plot")
```

Top most critics reviews are done by Glenn Kenny. Second largest critics review are done by AO Scott and Jeannette Catsoulis
