---
title: "Data-607 Week-2 Assignment"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 09/05/2018
* > [GitHub Location for rmd file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week2/SachidVijay.Deshmukh60.rmd)
* > [GitHub Location for SQL script](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week2/MovieRatings.sql)
* > [GitHub Location for pdf file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week2/SachidVijay.Deshmukh60.pdf)
* > [RPubs location of published file](http://rpubs.com/sachid/Data607-Week2Assignment)

##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Following are the pre-requisites for running this script
* Local installation of MySQL Server
* Local installation of MySQL Workbench
* MovieRatings.sql: Git Hub url is provided in the header section above
* R Packages RMySQL, htmlTable

#### Take following steps to execute the R script
* Download MovieRatings.sql file and run it against active schema on your MySQL server
* Make sure MovieRatings.sql script is executed successfully
* Open R mark up script in R Studio
* Change variables login.user, login.password and db.name variables to match your envronment
* Execute R mark up script

```{r reference library}
library(RMySQL)
library(htmlTable)
```

### 1] Configurable variables. Make sure to change the variables to match your environment

```{r declare variables}
login.user = 'sachid'
login.password = 'DatabasePassword@1'
db.name = 'moviecatalogue'

mydb = dbConnect(MySQL(), user= login.user, password=login.password, dbname=db.name)
```

### 2] Read Movie Ratings data into R Data Frame
```{r load data frame}
Movie.Ratings <- dbGetQuery(mydb, "select moviename, username, userrating from users join userratings 
                  on users.userid = userratings.userid join movies on userratings.movieid= movies.movieid order by movies.moviename")
```

### 3] Display movie ratings in  data frame. Displayed in matrix format
```{r display data frame matrix}
x = matrix(Movie.Ratings$userrating,ncol=5,byrow=T, dimnames = list(unique(Movie.Ratings$moviename), unique(Movie.Ratings$username)))
htmlTable(x, caption="Most populer movie ratings")
```

### 4] Display movie ratings in  data frame.
```{r display data frame}
htmlTable(Movie.Ratings, caption="Most populer movie ratings",rnames=F)
```