---
title: "Data-607 Week-1 Assignment"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 08/28/2018
##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1] Read Mushroom data from csv file

**Make sure to set header=F since data file don't have column names**

```{r Read File}
Mushroom.Data = read.csv("agaricus-lepiota.csv", header=F, stringsAsFactors=FALSE)
```

### 2] Assign column names

**We can use names function to assign column names**
**First column is the class label. Let's call it Edibility**

```{r Assign Columns}
names(Mushroom.Data) = c("Edibility", "CapShape", "CapSurface", "CapColor", "Bruises", "Odor", "GrillAttachment", "GrillSpacing", "GrillSize", "GrillColor", "StalkShape", "StalkRoot","StalkSurfaceAboveRing","StalkSurfaceBelowRing", "StalkColorAboveRing", "StalkColorBelowRing", "VeilType", "VeilColor", "RingNumber", "RingType", "SporeSprintColor", "Population", "Habitat")
```

### 3] Check Mushroom dataframe dimension. Rows and Columns

```{r check dimensions}
dim(Mushroom.Data)
```
**Mushroom dataframe has 8124 rows and 23 columns**

### 4] Check Edibility proportion

```{r check Edibility Proportion}
table(Mushroom.Data$Edibility)
```
**Out of 8124 samples 4208 samples are edible and 3916 samples are poisonous**

### 5] Subset columns
**Let's take first four attribute along with Edibility class variable

```{r Subset Columns}
Mushroon.Sub = subset(Mushroom.Data, select=c("Edibility", "CapShape", "CapSurface", "CapColor", "Bruises"))
```

### 6] Check dimensions of a subset dataframe

```{r Check Subset Dimensions}
dim(Mushroon.Sub)
```
**Subset has 8124 rows and 5 columns. First column is a class variable whihc is Edibility**

### 7] Preview first 10 rows of subset

```{r Preview Subset First}
head(Mushroon.Sub,10)
```
### 8] Preview last 10 rows of subset

```{r Preview Subset Last}
tail(Mushroon.Sub,10)
```

### 9] Replace abbreviations
**Replace e with Edible and p with Poisonous in class variable**

```{r Replace Abbreviations}
Mushroon.Sub$Edibility[Mushroon.Sub$Edibility=="e"] = "Edible"
Mushroon.Sub$Edibility[Mushroon.Sub$Edibility=="p"] = "Poisonous"
```
### 10] Check unique values of Class variable Edibility

```{r Validate Abbreviations}
unique(Mushroon.Sub$Edibility)
```
**We cas see that abbreaviatoins are successfully replaced**
