---
title: "Data-607 Week-4 Project1"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 09/23/2018
* > [GitHub Location for rmd file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week4/SachidVijay.Deshmukh60.rmd)
* > [GitHub Location for pdf file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week4/SachidVijay.Deshmukh60.pdf)
* > [RPubs location of published file](http://rpubs.com/sachid/Data607Week4Project1)
* > [Output file location file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week4/output.csv)



##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1] Library Initialization

```{r library initialization}
library(stringr)
```

### 2] Read Data
```{r Read Data}
col.names <- c("PairNum", "PalyerName", "TotPts", "Round1", "Round2", "Round3", "Round4", "Round5", "Round6", "Round7", "LastIndex")
player.data <- read.table("https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week4/tournamentinfo.txt", header = FALSE, skip = 4, sep = "|", fill = TRUE, stringsAsFactors = FALSE, col.names = col.names)[,1:10]
```

### 3] Remove white spaces
```{r Remove Padding}
player.data <- subset(player.data, !PalyerName == "")
player.data = as.data.frame(sapply(player.data, str_trim))
```
### 4] Add State and players pre rating columns
```{r add state and ranking}
player.tmp = player.data[seq(2, nrow(player.data), 2), ]
player.data =  player.data[seq(1, nrow(player.data), 2), ]
player.data$PlayerState = player.tmp$PairNum
player.data$PlayerPreRanking = str_trim(str_extract(player.tmp$PalyerName, "[[:blank:]]{1}[[:digit:]]{3,4}"))
```

### 5] Transform Opponent players
```{r transform points}
TransformPoints = function(x)
{
  str_trim(str_extract(x, "[[:space:]]+[[:digit:]]{1,2}"))
}
player.data[4:10] = sapply(player.data[4:10], TransformPoints)
```

### 6] Calculate opponent players pre-rating 
```{r Calc Prerating}
for (i in 1:nrow(player.data)) {
  for(j in 4:10) {
    player.data[i,j] <- player.data[player.data$PairNum == player.data[i,j],12][1]
  }
}

```

### 7] Calculate players avg rating
```{r Avg Rating}
player.data[4:10] = sapply(player.data[4:10], function(x) as.numeric(x))
player.data$AvgRating <- round(rowMeans(player.data[, c(4:10)], na.rm = TRUE), 1)
```

### 8] Create final output
```{r Final Output}
player.data.out = subset(player.data, TRUE, c(PalyerName, PlayerState, TotPts, PlayerPreRanking, AvgRating))
write.csv(player.data.out, file = "output.csv", row.names = FALSE)
```