---
title: "Data-607 Week-6 Assignment"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 10/04/2018
* > [GitHub Location for rmd file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week6/SachidVijay.Deshmukh60.rmd)
* > [GitHub Location for pdf file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week6/SachidVijay.Deshmukh60.pdf)
* > [RPubs location of published file](http://rpubs.com/sachid/Data607Week6Assignment)
* > [CSV file-1](https://raw.githubusercontent.com/mlforsachid/MSDSQ1/master/Data607/Week6/CrimeRate.csv)
* > [CSV file-2](https://raw.githubusercontent.com/mlforsachid/MSDSQ1/master/Data607/Week6/StudentGrades.csv)
* > [CSV file-3](https://raw.githubusercontent.com/mlforsachid/MSDSQ1/master/Data607/Week6/CityTemp.csv)

##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1] Library Initialization

```{r library initialization}
library(tidyr)
library(dplyr)
library(ggplot2)
```

### 2] Data Set-1 Analysis

#### Read Crime Rate data for three countries reported in year 1999

```{r Read Data}
Crime.Rate = read.csv("https://raw.githubusercontent.com/mlforsachid/MSDSQ1/master/Data607/Week6/CrimeRate.csv", stringsAsFactors = F)
```
#### Preview Data

```{r Preview Data}
head(Crime.Rate)
str(Crime.Rate)
```
#### Note that crimeinfo column is stacked. Cases indicates the crime cases reported for a specific country and population indicates total population of the country. Let's spread crimeinfo column

```{r spread column}
Crime.Rate = tidyr::spread(Crime.Rate, crimeinfo, value)
```

#### Preview unpivoted Crime Rate

```{r unpcrimerate}
head(Crime.Rate)
```

#### Note how crimerate column is spread based on categories. This also flattened the whole data frame.

### 3] Data Set-2 Analysis

#### Read Student Grade data for three subjects

```{r Read StudentGrData}
Stu.Grade = read.csv("https://raw.githubusercontent.com/mlforsachid/MSDSQ1/master/Data607/Week6/StudentGrades.csv", stringsAsFactors = F)
```
#### Preview Data

```{r Preview StudentGrData}
head(Stu.Grade)
str(Stu.Grade)
```

#### Note that each grades for a particular subject are on different column. Let's create a single Subject column.

```{r spread StudentGrcolumn}
Stu.Grade = tidyr::gather(Stu.Grade, "subject", "grades", 2:4)
```

#### Preview unpivoted Crime Rate

```{r StudentGrunpcrimerate}
head(Stu.Grade)
```

#### Note how individual columns for subject are collapsed into single column. The values are captured under newly creared grades column

### 4] Data Set-3 Analysis

#### Read City Temperature data for three cities

```{r Read CityTmpData}
City.Temp = read.csv("https://raw.githubusercontent.com/mlforsachid/MSDSQ1/master/Data607/Week6/CityTemp.csv", stringsAsFactors = F)
```
#### Preview Data

```{r Preview CityTmpData}
head(City.Temp)
str(City.Temp)
```

#### Note Date column. Let's separate Month, Day and Year into separate columns

```{r separate CityTmpcolumn}
City.Temp = tidyr::separate(City.Temp, "date", c("month", "day", "year"), sep="/")
```

#### Preview unpivoted Crime Rate

```{r CityTmpunpcrimerate}
head(City.Temp)
```

#### Note how date column is splitted across three separate columns (month, day and year)