---
title: "Data-607 Week-3 Assignment"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 09/16/2018
* > [GitHub Location for rmd file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week3/SachidVijay.Deshmukh60.rmd)
* > [GitHub Location for pdf file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data607/Week3/SachidVijay.Deshmukh60.pdf)
* > [RPubs location of published file](http://rpubs.com/sachid/Data607Week3Assignment)

##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library initialization}
library(stringr)
```
#### Question 3: Copy the introductory example. The vector name stores the extracted names.


```{r name vector}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name = unlist(str_extract_all(raw.data, "[[:alpha:] .,]{2,}"))
name
```

# * (a)
```{r first name, last name}
name.clean = str_replace_all(name,",", "")
name.fl = unlist(str_replace_all(name.clean, "[:alpha:]*[:punct:] ", ""))
name.fl
```
# * (b)

```{r has punct}
str_detect(name.clean, unlist(str_extract_all(name.clean, "[:alpha:]{2,}[:punct:]")))
```

# * (c)
```{r has second name}
str_detect(name.clean, unlist(str_extract_all(name.clean, "[A-Z][:punct:]{1}")))
```

#### Question 4: Describe the type of strings that conform to the following regular expression

# * (a) [0-9]+\\$ : This regex match one or more digit follwed by $ sign
# Example
```{r extract $ amount}
str = "The value of this product is 100$"
amount = unlist(str_extract_all(str, "[0-9]+\\$"))
amount
```
# * (b) \\b[a-z]{1,4}\\b : This regex match any word in lower case whihc is 1 to 4 chars long
# Example
```{r extract bat}
str = "This is my batbinton bat"
bat = unlist(str_extract_all(str, "\\b[a-z]{1,4}\\b"))
bat
```
# * (c) .*?\\.txt$ : This regex match any word endting wiht .txt
# Example
```{r .txt}
files = c("Program.R", "Program.cpp", "Program.txt")
txt = unlist(str_extract_all(files, ".*?\\.txt$"))
txt
```
# * (d) \\d{2}/\\d{2}/\\d{4} : This regex match date pattern mm/dd/yyyy
# Example
```{r Date Example}
str = "Today's date is 09/16/2018"
date = unlist(str_extract_all(str,"\\d{2}/\\d{2}/\\d{4}" ))
date
```

# * (e) <(.+?)>.+?</\\1> : This regex matches any word wrapped inside <> and </> HTML tags
# Example
```{r html extract}
str = "To print in bold write <b>Bold</b> in Html"
html = unlist(str_extract_all(str, "<(.+?)>.+?</\\1>"))
html
```
#### Question 9 : Secret Message
```{r secret message}
sm =  "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
decoded <- unlist(str_extract_all(sm, "[[:upper:].]{1,}"))
decoded <- str_replace_all(paste(decoded, collapse = ''), "[.]", " "); decoded
````