---
title: 'Data 606: Chapter-5 Homework'
output:
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

##---------------------------------------------------------------------------
###**Student Name :** Sachid Deshmukh
###**Date :** 11/16/2018
* > [GitHub Location for rmd file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data606/Week6/Homework5/SachidVijay.Deshmukh60.Rmd)
* > [GitHub Location for pdf file](https://github.com/mlforsachid/MSDSQ1/blob/master/Data606/Week6/Homework5/SachidVijay.Deshmukh60.pdf)
* > [RPubs location of published file](http://rpubs.com/sachid/Data606Homework5)

##---------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  > 5.6 Calculate the sample mean, the margin of error, and the sample standard deviation.

```{r statscalc}
N = 25
Margin.Error = (77-65)/2
Sample.Mean = 77 - Margin.Error
Crit.Value = qt(1-(0.05), (N-1))
Standard.Error = Margin.Error/Crit.Value
Sample.SD = Standard.Error * sqrt(N)
print(paste("Sample mean =", Sample.Mean))
print(paste("Margin of error =", Margin.Error))
print(paste("Sample standard deviation =", Sample.SD))
```

2. > 5.48 (A) Write hypotheses for evaluating whether the average number of hours worked varies across the groups.

Answer : 

Null Hypothesis : The means of each group are equal
Alternate Hypothesis : The means of each group are not equal

3. > 5.48 (B) Check conditions and describe any assumptions you must make to proceed with the test.

Answer : Sample sizes are fairly large. Data looks noramlly distributed. This is good for hypothesis testing. We need to assume independence across groups for achieving fair results in hypothesis testing

4. > 5.48 (C) Below is part of the output associated with this test. Fill in the empty cells.

```{r}
groups <- 5
samplesize <- 1172
meansq.group <- 501.54
sumsq.res <- 267382
alpha <- 0.0682

# Find Degree of Freedom
df.group <- groups-1
df.res <- samplesize - groups
df.total <- df.group + df.res

# Find Sum of Squared

sumsq.group <- df.group * meansq.group
sumsq.total <- sumsq.group + sumsq.res

# Find Mean Square
mean.square <- sumsq.res / df.res


# Find F-value
f.stat <- meansq.group / mean.square

print(paste("Degree of freedom for groups = ", df.group))
print(paste("Degree of freedom for residual = ", df.res))
print(paste("Degree of freedom for total = ", df.total))
print(paste("Sum of squared for groups = ", sumsq.group))
print(paste("Sum of squared for total = ", sumsq.total))
print(paste("Mean Square residual = ", mean.square))
print(paste("F Value = ", f.stat))
```

5. > 5.48 (D) What is the conclusion of the test

Answer - With 95% confidence interval, level of significance, alpha=0.05. Since p value=0.0682 > alpha, we fail to reject null hypothesis. We need to accept null hypothesis and assume that mean for each groups are equal

